Adaptive group-regularized logistic elastic net regression

### 2019-04-01 docs/manuscript.Rnw
added all suggestions

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,bbm,amsmath,amssymb,amsthm,
amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,
titlesec,multicol,verbatim,threeparttable,booktabs,dsfont,parskip,
xr-hyper,hyperref}
\externaldocument[sm-]{supplement}

@@

	Recently, \cite{zhang_novel_2019} introduced VBVS, a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
	adaptive inclusion of external information on the features.

	In line with the above, our proposed method is (a) subjective: the 
	hyper-parameters are estimated from the data and not chosen on objective 
	grounds, (b) adaptive: prior information is automatically and subjectively 
	weighed with respect to its informativeness, (c) fast compared to full 
	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
	user: only the data and grouping of the features is required. We emphasize that 
	our method is different from recent developments in grouped penalized 
	regression such as \cite{dondelinger_joint_2018}, where instead of the 
	features, the observations are partitioned.

@@

	\section{Simulations}\label{sec:simulations}
	To assess classification and variable selection performance we conducted a 
	simulation study in which we compared \texttt{gren} to the regular elastic 
	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, 
	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
	composite mimimax concave penalty (\texttt{cMCP}) 
	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
	\texttt{gren} in the sense that it estimates group-specific penalty 
	multipliers. The two main differences with \texttt{gren} are (i) the absence 
	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
	are extensions of the group lasso and not adaptive on the group level. 
	However, in contrast to the original group lasso,
	they select single features, instead of just whole groups.

	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{cMCP}, 
	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
	such that the first setting closely resembles the ridge setting, 
	where $\alpha=0$, while the third setting is similar to the lasso with 
	$\alpha=1$. In \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}, $\alpha$
	determines the ratio of group-level penalization ($\alpha=0$) to 
	feature level ($\alpha=1$) penalization. For the methods that do 
	automatic feature selection, we estimate models for a range of model sizes:
	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model
		parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations 
		within groups of features,
		\item differential signal between the groups and heavy-tailed distributed 
		model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 
	\ref{sm-sec:simulations} of 
	the SM. In the first scenario, the distribution of model parameters is far
	from the assumed ridge/elastic net/lasso distributions, so we expect
	all methods to underperform. In the second scenario, which is similar to
	the simulations presented in the original group lasso paper 
	\cite[]{yuan_model_2006}, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate. The group lasso extensions are expected to perform better: they 
	estimate just one penalty parameter, so are less sensitive to overfitting.
	In scenario three, we expect all methods to underperform
	due to the high correlations between the features. The fourth 
	scenario is expected to yield a better \texttt{gren} performance, since
	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
	detect differential signal between
	the groups. Scenario five is expected to yield a bad performance for all 
	methods, because of the sparse truth and lack of differential 
	signal.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples to compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
	$p=900$ in scenario 2 and $p=1000$ in the other scenarios. We varied the 
	covariance matrix of the features and simulation of 
	the model parameters $\beta_j$, $j=1, \dots, p$ between the
	scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
	parametrisations). We 
	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation, we repeated each scenario 100 times 
	and report predictive performance measures 
	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
	Figures \ref{fig:lines_simulations_res1_auc_briers}--[! figure refs here]. 
	Additional results and performance measures are presented in Section 
	\ref{sm-sec:simulations} of the SM.

	- description of results

@@

@@

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
  three methods, we only present \texttt{cMCP} here; the results for 
  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
  [! section ref here]. For the methods with a
  tuning parameter $\alpha$, we picked the $\alpha$ 
  (two $\alpha$'s for \texttt{gren}) that gave the best
  performance and refer
  the reader to SM Section \ref{sm-sec:colorectal} for comparisons with 
  different $\alpha$'s.

@@

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes in terms of AUC 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
  This is especially pronounced for the larger models. The performance of
  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
  all group lasso extensions that we investigated.
  (SM Section \ref{sm-sec:colorectal}). Overall, the 
  inclusion of the extra information on the 
  features benefited predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit it only slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@

@@

  We compared \texttt{gren} to the same methods as in Section 
  \ref{sec:colorectal}. However, since the feature information
  consists of two partitions with overlapping groups, we used an extension of
  \texttt{cMCP} that allows for
  overlapping groups \cite[]{zeng_overlapping_2016}.
  We estimated AUC for the methods on an independent validation set of 97 
  patients \cite[]{mes_prognostic_2017}. The features 
  in this validation set were
  microarray expression levels, normalised to account
  for a difference in scale compared to the RNAseq training data. We present the
  results on this validation set, together with the estimated penalty 
  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Additional
  comparisons and performance measures are presented in the SM Section 
  \ref{sm-sec:oral}.

  The estimated penalty multipliers follow the expected pattern 
  (Figure \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}a): 
  larger cis-correlations receive smaller penalties. 
  Likewise, smaller external $p$-values are penalized less. This pattern is
  slightly less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the
  other methods. The predictive performance of
  \texttt{gren} is slighty better than the other methods for a range of model
  sizes. In this example, ridge and \texttt{GRridge} perform almost 
  identical, while \texttt{gren} outperforms the regular elastic net.
<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the lymph node metastasis example.", out.width="100%", fig.asp=1>>=
@

  \section{Additional applications}\label{sec:applications}
  Two more applications are presented in SM Sections \ref{sm-sec:alzheimer} and
  \ref{sm-sec:cervical}. The first one is concerned with diagnosis of 
  Alzheimer's from 230 metabolites' expression levels in 174 subjects. We 
  included several sources of extra information on the metabolites. In this 
  example, \texttt{gren} performs worse than the group lasso methods. This is 
  due to the smaller number of features to learn the penalty parameters from. 
  Additionally, the metabolites are highly collinear, which further impairs 
  estimation. The group lasso extensions depend on just one penalty parameter,
  so suffer less from the small number of features.

@@

	Although our method is essentially a reweighted elastic net and can be
	considered weakly adaptive, it is different from 
	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
	rather than to the primary data. It also differs in the scale of adaptation: 
	in the adaptive lasso and elastic net the adaptive weights are feature 
	specific, while in our case they are estimated on the group level, rendering
	the adaptation more robust. As can be seen from both the simulations in 
	Section \ref{sec:simulations} and the applications to different outcomes 
	(prognostic or diagnostic) and marker types 
	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
	\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}--
	\ref{sm-sec:cervical}, adaptation to 
	external data may be beneficial for prediction and feature selection. 
	We believe that this is due to the `borrowing of information' effect: 
	estimates of feature effects that behave similarly are shrunken similarly, 
	yielding overall, better estimates.

@@

bench2 <- read.table("results/rnaseq_lymph_node_metastasis_bench1.csv", 
                     stringsAsFactors=FALSE)
times1 <- round(bench1$time*10^(-9), 2)
times2 <- round(bench2$time*10^(-9), 2)
session <- gsub("_", ".", devtools::session_info()$platform$system)
@
  The double EM loop runs the danger of excessive computation time. In our 
  implementation we have reduced computational time considerably by applying
  some computational tricks (see SM Section 
  \ref{sm-sec:computationalcomplexity}). In addition, the computationally heavy
  parts were implemented in \texttt{C++} to further reduce the computation time. 
  To assess this reduction, we have compared computation times to the
  other methods introduced in Sections \ref{sec:colorectal}--
  \ref{sec:oral}. All timings were obtained on a Macbook 
  Pro 2016 running \Sexpr{session}. 
  The timings of \texttt{gren} are competitive. On the colorectal cancer example
  we achieved the following times:
  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
  respectively. \texttt{GRridge} and \texttt{SGL} with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ are 
  slightly faster with \Sexpr{times1[4]} and \Sexpr{times1[c(5:7)]} seconds, 
  respectively.
  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
  The results for the lymph node cancer example are similar. \texttt{gren} 
  achieved 
  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
  respectively.
  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}. 
  % We note that a considerable speedup may be achieved by iterating between 
  % single EB and single VB updates, instead of running the VB EM 
  % until convergence. This works in practice, because an EM tends to take a few 
  % initial large steps towards the optimum, followed by a series of smaller 
  % steps. We argue that the first few iterations of the VB EM render the 
  % approximate expectation in \ref{eq:approximatedloglikelihood}
  % accurate enough for a 
  % reliable EB step.

@@

	\section*{Supplementary Material}
	Supplementary Material is available online from 
	\url{http://pub.math.leidenuniv.nl/~munchmm/}. The results and documents may 
	be recreated from \url{https://github.com/magnusmunch/gren}.

>>>>

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,bbm,amsmath,amssymb,amsthm,
amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,
titlesec,multicol,verbatim,threeparttable,booktabs,dsfont,parskip,
xr-hyper,hyperref,enumerate}
\externaldocument[sm-]{supplement}

@@

	In line with the above, our proposed method is (a) objective: the 
	hyper-parameters are estimated from the data and not chosen on subjective 
	grounds, (b) adaptive: prior information is automatically and subjectively 
	weighed with respect to its informativeness, (c) fast compared to full 
	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
	user: only the data and grouping of the features is required. We emphasize 
	that our method is different from recent developments in grouped penalized 
	regression such as \cite{dondelinger_joint_2018}, where instead of the 
	features, the observations are partitioned.

	Recently, \cite{zhang_novel_2019} introduced Variational Bayes Variable
	Selection (VBVS), a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
	adaptive inclusion of external information on the features.

@@

	\section{Simulations}\label{sec:simulations}
	To assess classification and variable selection performance we conducted a 
	simulation study in which we compared \texttt{gren} to the regular elastic 
	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016},     
	composite mimimax concave penalty (\texttt{cMCP}) 
	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
	\texttt{gren} in the sense that it estimates group-specific penalty 
	multipliers. The two main differences with \texttt{gren} are (i) the absence 
	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
	are extensions of the group lasso and not adaptive on the group level. 
	However, in contrast to the original group lasso,
	they select single features, instead of complete groups.

	For the regular elastic net, \texttt{gren}, \texttt{cMCP}, 
	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
	such that the first setting closely resembles the ridge setting, 
	where $\alpha=0$, while the third setting is similar to the lasso with 
	$\alpha=1$. In \texttt{cMCP} and \texttt{gel}, $\alpha$
	determines the ratio of group-level penalization ($\alpha=0$) to 
	feature level ($\alpha=1$) penalization. For the methods that do 
	automatic feature selection, we estimate models for a range of model sizes:
	$\hat{p} \in \{2, 4, 8, 16, 32, 64, 128, 256\}$.

	We simulated data according to five different scenarios:
	\begin{enumerate}[(i)]
		\item \label{it:scenario1} differential signal between the groups and uniformly
		distributed model
		parameters,
		\item \label{it:scenario2} a large number of small groups of features,
		\item \label{it:scenario3} no differential signal between the groups, but 
		strong correlations 
		within groups of features,
		\item \label{it:scenario4} differential signal between the groups and 
		heavy-tailed distributed 
		model parameters, and
		\item \label{it:scenario5}a very sparse setting with no signal in some of 
		the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 
	\ref{sm-sec:simulations} of 
	the SM. In Scenario (\ref{it:scenario1}), the distribution of model 
	parameters is far
	from the assumed ridge/elastic net/lasso distributions, so we expect
	all methods to underperform. In Scenario (\ref{it:scenario2}), 
	which is similar to
	the simulations presented in the original group lasso paper 
	\cite[]{yuan_model_2006}, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate. The group lasso extensions are expected to perform better: they 
	estimate just one penalty parameter, so are less sensitive to overfitting.
	In Scenario (\ref{it:scenario3}), we expect all methods to underperform
	due to the high correlations between the features. The
	Scenario (\ref{it:scenario4}) is expected to yield a better \texttt{gren}
	performance, since
	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
	detect differential signal between
	the groups. Scenario (\ref{it:scenario5}) is expected to yield a bad 
	performance for all 
	methods, because of the sparse truth and lack of differential 
	signal.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples to compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where $p$ varies 
	from $p=900$ to $p=1000$ over the Scenarios. We varied the 
	covariance matrix of the features and simulation of 
	the model parameters $\beta_j$, $j=1, \dots, p$ between the
  Scenarios (see SM Section \ref{sm-sec:simulations} for the exact 
	parametrisations). We 
	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation, we repeated each scenario 100 times 
	and report predictive performance measures 
	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
	Figures \ref{fig:lines_simulations_res1_auc_briers}--[! figure refs here]. 
	For the methods with an $\alpha$ tuning parameter, we show the one that gives
	the best performance. Results for the other $\alpha$'s and additional
	performance measures are presented in Section 
	\ref{sm-sec:simulations} of the SM.

	In Scenario (\ref{it:scenario1})
	\texttt{gren} and \texttt{GRridge} don't suffer much from the wrong 
	distributional assumptions as they outperform the other methods.
	Scenario (\ref{it:scenario2}) follows our expectation \texttt{gren}, and to
	a lesser extent \texttt{GRridge} underperform due to overfitting. Suprisingly,
	the regular elastic net outperforms the group lasso methods, which are 
	designed to handle situations such as Scenario (\ref{it:scenario2}).
	In Scenario (\ref{it:scenario3}), \texttt{gren} and to a lesses extent the
	regular elastic net suffer more from the higher correlations. Here,
	the regular elastic net outperforms \texttt{gren}, which is an indication
	that the estimation of the penalty parameters is impaired by high
	correlation between features. Scenario (\ref{it:scenario4}) follows the 
	expected pattern, with the group-regularized methods (\texttt{GRridge} and
	\texttt{gren}) outperforming their respective non group-regularized 
	counterparts, as well as the sparse group lasso extensions.
	Surprisingly, in Scenario (\ref{it:scenario5}), \texttt{gren} outperforms all
	other methods, which is an indication that \texttt{gren} is able to pick up 
	very sparse differential signals between the groups.

@@

<<lines_simulations_res2_auc_briers, fig.cap="Estimated (a) AUC and (b) Brier skill score in simulation setting 2.", out.width="100%", fig.asp=1/2>>=
@

@@

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
  three methods, we only present \texttt{cMCP} here; the results for 
  another extension of the group lasso, the sparse
  group lasso by \cite{simon_sparse-group_2013}
  (\texttt{SGL}), and \texttt{gel} are presented in the SM Section 
  \ref{sm-sec:colorectal}. For the methods with a
  tuning parameter $\alpha$, we picked the $\alpha$ 
  that gave the best performance and refer
  the reader to SM Section \ref{sm-sec:colorectal} for comparisons with 
  different $\alpha$'s.

@@

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
  \texttt{gren} seems to outperform the other methods for a range of smaller 
  model sizes in terms of AUC 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
  Selection of larger models is difficult with large values of $\alpha$, an 
  inherent property of the strong $L_1$-norm penalization. With lower values of
  $\alpha$, larger models are possible (see SM Section \ref{sm-sec:colorectal})
  for comparisons with smaller values of $\alpha$). The performance of
  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
  all group lasso extensions that we investigated.
  (SM Section \ref{sm-sec:colorectal}). Overall, the 
  inclusion of the extra information on the 
  features benefited predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit only slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@

@@

  We compared \texttt{gren} to the same methods as in Section 
  \ref{sec:colorectal}. However, since the feature information
  consists of two partitions with overlapping groups, we used an extension of
  \texttt{cMCP} that allows for
  overlapping groups \cite[]{zeng_overlapping_2016}.
  We estimated AUC for the methods on an independent validation set of 97 
  patients \cite[]{mes_prognostic_2017}. The features 
  in this validation set were
  microarray expression levels, normalised to account
  for a difference in scale compared to the RNAseq training data. We present the
  results on this validation set, together with the estimated penalty 
  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
  \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}. Additional
  comparisons and performance measures are presented in the SM Section 
  \ref{sm-sec:oral}.

  The estimated penalty multipliers follow the expected pattern 
  (Figure \ref{fig:barplot_lines_rnaseq_oral_cancer_res1_auc}a): 
  larger cis-correlations receive smaller penalties. 
  Likewise, smaller external $p$-values are penalized less. This pattern is
  slightly less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the
  other methods. The predictive performance of
  \texttt{gren} is slighty better than the other methods for a range of model
  sizes. In this example, ridge and \texttt{GRridge} perform almost 
  identical, while \texttt{gren} outperforms the regular elastic net.
<<barplot_lines_rnaseq_oral_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the oral cancer example.", out.width="100%", fig.asp=1>>=
@

  \section{Additional applications}\label{sec:applications}
  Two more applications are presented in SM Sections \ref{sm-sec:alzheimer} and
  \ref{sm-sec:cervical}. The first one is concerned with diagnosis of 
  Alzheimer's from 230 metabolites' expression levels in 174 subjects. We 
  included several sources of extra information on the metabolites. In this 
  example, \texttt{gren} performs worse than the group lasso methods. This is 
  due to the smaller number of features to learn the penalty parameters from. 
  Additionally, many metabolites are strongly negatively correlated, which
  further impairs estimation. The group lasso extensions depend on just one
  penalty parameter, so suffer less from the small number of features and
  negative correlations.

@@

	Although our method is essentially a reweighted elastic net and can be
	considered weakly adaptive, it is different from 
	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
	rather than to the primary data. It also differs in the scale of adaptation: 
	in the adaptive lasso and elastic net the adaptive weights are feature 
	specific, while in our case they are estimated on the group level, rendering
	the adaptation more robust. As can be seen from both the simulations in 
	Section \ref{sec:simulations} and the applications in Sections 
	\ref{sec:colorectal}-\ref{sec:oral}, and SM Sections \ref{sm-sec:colorectal}-
	\ref{sm-sec:cervical} adaptation to 
	external data may be beneficial for prediction and feature selection. 
	We show that to be true for both prognostic and diagnostic targets,
	as well as a range of marker types (RNAseq, microRNAs, or metabolomics).
	We believe that this is due to the `borrowing of information' effect: 
	estimates of feature effects that behave similarly are shrunken similarly, 
	yielding overall, better estimates.

@@

bench2 <- read.table("results/rnaseq_oral_cancer_bench1.csv", 
                     stringsAsFactors=FALSE)
times1 <- round(bench1$time*10^(-9), 2)
times2 <- round(bench2$time*10^(-9), 2)
session <- gsub("_", ".", devtools::session_info()$platform$system)
@
  The double EM loop runs the danger of excessive computation time. In our 
  implementation we have reduced computational time considerably by applying
  some computational tricks (see SM Section 
  \ref{sm-sec:computationalcomplexity}). In addition, the computationally heavy
  parts were implemented in \texttt{C++} to further reduce the computation time. 
  To assess this reduction, we have compared computation times to the
  other methods introduced in Sections \ref{sec:colorectal}--
  \ref{sec:oral}. All timings were obtained on a Macbook 
  Pro 2016 running \Sexpr{session}. 
  The timings of \texttt{gren} are competitive. On the colorectal cancer example
  we achieved the following times:
  \Sexpr{times1[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
  respectively. \texttt{GRridge} and \texttt{SGL} with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$ are 
  slightly faster with \Sexpr{times1[4]} and \Sexpr{times1[c(5:7)]} seconds, 
  respectively.
  \texttt{cMCP} (\Sexpr{times1[c(8:10)]} seconds) and \texttt{gel} 
  (\Sexpr{times1[c(11:13)]} seconds) are faster than \texttt{gren}. 
  The results for the oral cancer example are similar. \texttt{gren} 
  achieved 
  \Sexpr{times2[c(1:3)]} seconds, with $\alpha \in \{ 0.05, 0.5, 0.95 \}$, 
  respectively.
  \texttt{GRridge} was slightly faster with \Sexpr{times2[4]} seconds, 
  while \texttt{SGL} was slightly slower: \Sexpr{times2[c(5:7)]} seconds, with 
  $\alpha \in \{ 0.05, 0.5, 0.95 \}$.
  \texttt{cMCP} (\Sexpr{times2[c(14:16)]} seconds) and \texttt{gel} 
  (\Sexpr{times2[c(17:19)]} seconds) are again faster than \texttt{gren}. 
  % We note that a considerable speedup may be achieved by iterating between 
  % single EB and single VB updates, instead of running the VB EM 
  % until convergence. This works in practice, because an EM tends to take a few 
  % initial large steps towards the optimum, followed by a series of smaller 
  % steps. We argue that the first few iterations of the VB EM render the 
  % approximate expectation in \ref{eq:approximatedloglikelihood}
  % accurate enough for a 
  % reliable EB step.

@@

	\section*{Supplementary Material}
	Supplementary Material is available online from 
	\url{http://biostatistics.oxfordjournals.org}.

	\section*{Reproducible Research}
	All results and documents may be recreated from 
	\url{https://github.com/magnusmunch/gren}.
